Zero-shot adaptation by meta-mapping between tasks

### 2019-09-05 writing/embedded_meta_learning_main.tex
Incorporating some of Jay's added comments

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recurse on its representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. We suggest that approaches like ourswill be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), we can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
There are different ways this adaptation can be cued, for example by examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for individual data points and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{subfigure}[b]{0.72\textwidth}
\begin{tikzpicture}[auto, scale=0.66, every node/.style={scale=0.85}]
\node [dashblock] at (0.2, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: Shared \\ representation\end{tabular}};

@@

More formally, we take a \emph{functional} perspective on learning. A datum can be represented by a constant function which outputs that datum. (For example, each point in the latent space of an autoencoder can be thought of this way.) Adopting this perspective allows us to interpret model inputs or outputs as functions. Most machine learning tasks then become a mapping of functions to functions. The key insight is that, while these mappings could be over functions that represent data\footnote{Where ``data'' is a quite flexible term. The approach is agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.}, they could also operate on functions that represent tasks. This is related to the idea of a homoiconicity, defined above. Under this perspective, basic tasks and meta-mappings are really the same type of problem. \par

>>>>

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recursively treat its task representations as data points that can be transformed to produce new task representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. We suggest that approaches like ours will be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
There are different ways this adaptation can be cued, for example by examples of similar task mappings (``you've played winning and losing poker, now try to lose at blackjack'') or simply by natural-language instructions (``try to lose at blackjack'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system that embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for individual data points and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system can learn to perform meta-mappings to follow instructions or change behavior. That is, it can transform task representations using the same components it uses to transform basic data points, because the task representations and basic data points are embedded in the same space. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, as we will show, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{subfigure}[b]{0.72\textwidth}
\begin{tikzpicture}[auto, scale=0.66, every node/.style={scale=0.85}]
\node [dashblock] at (0.2, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: Shared \\ representation\end{tabular}};

@@

More formally, we take a \emph{functional} perspective on learning. From this perspective, the data points that one function receives can themselves be functions\footnote{Indeed, any data point can be represented as a constant function that outputs the data point.}. The key insight is that then our architecture can transform data points\footnote{Where ``data'' is a quite flexible term. The approach is agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.} to perform basic tasks (as is standard in machine learning), but it can also transform these task functions to adapt to new tasks. This is related to the idea of a homoiconicity, defined above. Under this perspective, basic tasks and meta-mappings from task to task are really the same type of problem. The functions at one level of abstraction (the basic tasks) become inputs and outputs for higher-level functions at the next level of abstraction (meta-mapping between tasks). \par
